Determinism, Complexity, and Predictability in Computer Performance

### 2014-02-16 pecomplexity.tex
new edits

\begin{abstract}
  Computers are deterministic dynamical systems \cite{mytkowicz09}.
  Among other things, that implies that one should be able to use
  deterministic forecast rules to predict aspects of their behavior.
  That statement is sometimes---but not always---true. The memory and
  processor loads of some simple programs are easy to predict, for
  example, but those of more-complex programs like {\tt gcc} are not.
%%%%%%%%%%%%%%%%%%%%%%
%% I had to change all the \verb|blah| entries because they caused
%% latex to barf if they were in figure captions.  Odd bug.
%%%%%%%%%%%%%%%%%%%%%%
  The goal of this paper is to determine why that is the case. We
  conjecture that, in practice, complexity can effectively overwhelm
  the predictive power of deterministic forecast models. To explore
  that, we build models of a number of performance traces from
  different programs running on different Intel-based computers. We
  then calculate the \emph{permutation entropy}---a temporal entropy
  metric that uses ordinal analysis---of those traces and correlate
  those values against the prediction success.
\end{abstract}

@@

BRAINSTORMING 
\begin{itemize}
\item Complexity need not be hard to predict (can point at the simple predictions paper) [[move to introduction]]
\item random walk for example is best predicted by guess what just happend[[move to introduction]]
\item The kind of complexity present matters, i.e., that is whether the complexity is structured or not.[[use here and mention in intro]] 
\item Quantifying structured and unstructured complexity is nontrivial in the case of real-valued noisy time series but WPE does this. [[talk about again here but justify in information theory section]]
%\item Maybe plot a big chunk of \col and a big chunk of \gcc together and show that they both look complex.

\item \gcc appears visually very complex, *and* according to WPE this complexity is unstructured. And a constant, linear and nonlinear prediction strategy all fail. We should be able to conclude that guessing random values is the best we can do as is shown by MASE [[Use in this section]]

\item \col is also complex (can even be chaotic/ point to CHAOS paper) but the complexity is structured according to WPE and as such that complexity is usable for prediction [[use in this section]]

@@

THINGS TO DO
\begin{enumerate}
\item rewrite for consistency in notation. For example, some parts of the paper call word length $l$ and some call it $n$.
\item make sure it is clear and we justify that detecting and characterizing structural complexity is hard for real-valued noisy time series.
\item justify and explain the regime split in \svd 
\end{enumerate}

@@

[[Joshua: These need to be rewritten to conclude and address this paper...]]

>>>>

\begin{abstract}
[[Joshua: Needs to be rewritten to be more general and use computers as the application not the focus]]  Computers are deterministic dynamical systems \cite{mytkowicz09}.
  Among other things, that implies that one should be able to use
  deterministic forecast rules to predict aspects of their behavior.
  That statement is sometimes---but not always---true. The memory and
  processor loads of some simple programs are easy to predict, for
  example, but those of more-complex programs like \gcc are not.
%%%%%%%%%%%%%%%%%%%%%%
%% I had to change all the \verb|blah| entries because they caused
%% latex to barf if they were in figure captions.  Odd bug.
%%%%%%%%%%%%%%%%%%%%%%
  The goal of this paper is to determine why that is the case. We
  conjecture that, in practice, complexity can effectively overwhelm
  the predictive power of deterministic forecast models. To explore
  that, we build models of a number of performance traces from
  different programs running on different Intel-based computers. We
  then calculate the \emph{permutation entropy}---a temporal entropy
  metric that uses ordinal analysis---of those traces and correlate
  those values against the prediction success.
\end{abstract}

@@

BRAINSTORMING 
\begin{itemize}
\item Complexity need not be hard to predict (can point at the simple predictions paper) [[move to introduction]]
\item random walk for example is best predicted by guess what just happend[[move to introduction]]
\item The kind of complexity present matters, i.e., that is whether the complexity is structured or not.[[use here and mention in intro]]

@@

THINGS TO DO
\begin{enumerate}
\item rewrite for consistency in notation. For example, some parts of the paper call word length $l$ and some call it $n$.

\item make sure it is clear and we justify that detecting and characterizing structural complexity is hard for real-valued noisy time series.
\item Quantifying structured and unstructured complexity is nontrivial in the case of real-valued noisy time series but WPE does this. [[talk about again here but justify in information theory section]]
\item justify and explain the regime split in \svd 
\end{enumerate}

@@

\begin{enumerate}
\item  This need to be rewritten to conclude and address this paper...
\end{enumerate}
